The Occurrence of Earth-Like Planets Around Other Stars

### 2015-03-11 paper/response.txt

For the editor
--------------

As you will see from our response to the referee attached, we have
identified a number of errors in the referee's discussion of our work.
Additionally, the referee seems unusually hostile to our modelling
choices (e.g. "I could go on with many more effects that are not
adequately modeled, and often not even discussed. But it's not a
referee's job to develop a research plan for them."). As you will see
in the response, we feel that our choices are well-justified.

Given these issues, could we request to have a second, independent
referee comment on our revised work? If appropriate we would be happy
for both referees to see this note to the editor and the attached
referee response.

@@

We have made some changes to the discussion in the paper in response
to the current referee's misunderstandings; those are outlined in the
course of the response.

Response
--------

The referee has made several valid criticisms of our paper, which we
address below. Nevertheless, we feel that there are some
misunderstandings in the referee's main points of criticism. Below we
explain the changes we have made to our manuscript to attempt to
address the referee's criticism and where we feel that criticism is in
error.

@@

It is simply incorrect to say that our model is only validated by
comparison to candidates at much larger than Earth radius and shorter
than one-year periods. The referee *is* correct that there are
relatively few candidates in the P-R plane near 1 year and 1 REarth,
but this is exactly why it is important to validate the model in a
wide range of periods and radii that span the 1 year - 1 REarth region
of interest. We admit to being surprised that such a simple model
reproduced the data so well, but there it is.

The referee writes:

  Further, the manuscript assumes that the apparent size-period
  distribution of false positives can be described by a single
  two-dimensional normal distributions and suggest that this is
  consistent with background eclipsing binary stars being the dominant
  source of false positives among Kepler planet candidates. However,
  previous works have demonstrated that both hierarchical triple stars
  including an eclipsing binary and large planets orbiting the fainter
  member of a binary star are both significant sources of false
  positives. Further, both the rate of false positives and the relative
  abundance of the types of false positives are functions of orbital
  period and planet candidate size. This analysis completely ignores
  these important issues.

First, our manuscript describes the size-period distribution of false
positives by a linearly-varying function of size and period (c.f. Eq.
(6)). Second, in the third sentence the referee seems to be
complaining that we do not allow for size and period variations in the
background distribution, but it is explicit in both the text and
Figure 1 that the distribution of background contamination we use
varies over the size-period plane. Third, in the final sentence the
referee claims that we ignore these issues (related to contamination),
but we do not; for example, we note in the Conclusion with respect to
Petigura et al (2013) and Foreman-Mackey et al (2014) that "[in
contrast to our work] Neither study attempted to account for
contamination from falsely-identified candidate transit events,
controlling this instead through careful choice of threshold." Similar
statements appear throughout the paper. It is our view that
incorporating contamination in the analysis is an important advance
from our paper. The referee notes that many types of contamination are
present in the data, whose relative importance varies across the
size-period plane; we agree, but note that our contamination model is
purely empirical and agnostic to the source of contamination. While we
claim that our fit is consistent with the amount of contamination from
background eclipsing binaries determined from careful studies, nothing
in our model is based on the detailed properties of this particular
source of contamination. We have added several sentences throughout
the paper that emphasise this point. Finally, we have become aware
though helpful comments from Tim Morton that Morton & Johnson (2011)
presents an early measurement of contamination in Kepler, and we now
cite that paper in addition to Fressin et al. (2013) where
appropriate.

@@

The referee writes:

  The analysis is based entirely on a publicly available planet
  candidate list, which is heterogeneous, but these issues are not
  addressed. Further, they do not provide enough information about the
  data set used to allow someone else to repeat their calculation.

We state in the introduction that "The Q17 data release describes 2598
"candidate" planetary transit signals identified by the Kepler team
from observations of stars in the "EX" observing program (which are
primarily G-type main-sequence stars similar to our own Sun (Batalha
et al. 2010)), giving the inferred planetary period and radius for
each." and, following Eq. (1) discussing our fiducial SNR, that "In
our analysis, we obtain these quantities from the Kepler Input Catalog
(Batalha et al. 2010; Brown et al. 2011) and the MAST Kepler
archive[web link]." Additionally, we have released all our code for
the analysis at https://github.com/farr/kepler-selection (linked in
the paper's conclusion). This includes the code that processes the
data from the above-mentioned sources, code that performs the fitting,
and code that produces the plots found in the paper. We are not sure
what additional information the referee would require to repeat our
calculation. We also note that, to our knowledge, no other estimate of
eta-Earth has been performed using this data set (the result from
Petigura et al. 2013 and the following Foreman-Mackey et al. 2014 used
the some of the same lightcurves from Kepler, but a customised
pipeline). In some senses, this data set is heterogeneous (F, G, K
stars), but in many important ways it is not: same instrument
(Kepler), same pipeline, etc.

@@

The referee writes:

  I could go on with many more effects that are not adequately
  modeled, and often not even discussed. But it's not a referee's job
  to develop a research plan for them.

  In summary, the manuscript presents one basic calculation for an
  interesting, but complex problem which deserves a more detailed
  analysis. I can understand journals accepting one or two early
  analyses that are overly simplistic. But by now the field has plenty
  of simplistic estimates. To make a significant contribution, the field
  needs people to devote the time required to perform a more complete
  analysis that pays attention to both astronomical and statistical
  issues. Such a paper is almost certainly longer than appropriate for
  ApJL. But there is certainly room for at least several more analyses
  that build on each other to develop a progressively more complete
  model to understand the frequency and distribution of small planets
  around other stars.

We disagree with the referee that our calculation is "basic" or
"overly simplistic." Our approach to incorporating contamination and
emprically estimating selection effects is based on novel statistical
methods and represents a step forward relative to previous analyses.
The application of these methods to a previously-unexamined data set
provides a useful independent estimate of the size-period distribution
of exoplanets that confirms some aspects of previous analyses and
calls others into question. The fact that our simple model of the
distribution and selection functions fits the data extremely well
indicates that, at present, the data quality and quantity are not
sufficient to warrant more complicated modeling.
